---
title: "Descriptive R nOtebook"
output: html_notebook
---

libraries
```{r}
library(dplyr)
library(ggplot2)
library(corrplot)

```
load data
```{r}
data<- read.csv( "C:/Forscher/3891-2019/Arbeitsverzeichnis/total14_17.csv")
data<- data[, -c(1)]


```

missing varaibles -out 
```{r}
p<- function(x){sum(is.na(x))/length(x)*100}
apply(data, 2, p)
```

```{r}
#print number of missing values per element 

missing_ct<-colSums(apply(X =data, MARGIN = 2, is.na)) 
missing_ct[which(missing_ct > 0)]
print(paste("There are ", length(missing_ct[which(missing_ct > 0)]), 
            " variables have NAs.")) 
print(paste("Total number of NAs:", sum(missing_ct)))
```

Delete variables with more than 75 unknowns in each columns

```{r}
data<-data[,-c(23,27)]
```

```{r}
# as factors
colzz <- c(2:5,10:24,26,28,33)

data[, colzz]<-lapply(data[,colzz], factor)
data$na_count<- apply(is.na(data), 1, sum)
data$percent_na<-apply(data, 1, function(x){sum(is.na(x))/length(x)*100})

```

distribution of nas by rows 
```{r}
na_distribution <- ggplot(data, aes(na_count)) + geom_histogram(position = "dodge") 
na_distribution
```

delete rows with more than 75% nas, that is minimum 12 nas excluding the goe varaibles which were not used inmodelling 
```{r}
#total<-data%>%
 # filter(na_count <= 13) #75% of 50, 583508 obsve of 50 variables
total <-data #1849469 with 50 vars 
```

acdidents involving only personal injury 

```{r}
total <-total %>%
  filter(UKATEGORIE %in% 1:3) #1849469 obs and 56 vars 
```




Descriptive analysis by unovariate 
```{r}
by_ruralUrban <- total%>% select(UID,UORTSLAGE)%>% group_by(UORTSLAGE) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)


by_ruralUrban
# chart1
chartruralurban <- ggplot(data=by_ruralUrban, aes(x=UORTSLAGE, y=total.count, fill=UORTSLAGE))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Urban_or_Rural_Area")
chartruralurban
```


Number of accident per year 

```{r}
#by-year 
by_year_count <- total%>% select(UID,UJAHR)%>% group_by(UJAHR) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)

by_year_count 
# chart 1
chartyear <- ggplot(data=by_year_count, aes(x=UJAHR, y=total.count, fill=year))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Year")+geom_text(aes(label=paste0(round(total.count, 2),"%")),position=position_dodge(width=0.9), vjust=-0.25)

chartyear
```

accident severity per year 
```{r}
by_yearsev <- data %>% select(UID,UJAHR,UKATEGORIE) %>% group_by(UJAHR,UKATEGORIE) %>% summarise(total.count=n())%>% mutate(total.count = total.count/sum(total.count)*100)
by_yearsev

chart_yearsev <- ggplot(data=by_yearsev, aes(x=UJAHR, y=total.count, fill=UKATEGORIE)) + geom_bar(stat="identity")+geom_text(aes(label=paste0(round(total.count, 2),"%")),position=position_stack(vjust=0.25))+labs(y="Percentage", x= "Year")+ scale_fill_discrete(name= "Accident Severity", labels=c("Slight", "Serious", "Fatal"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())

chart_yearsev

```

Accident severity
```{r}
#Accident severity 

by_Accidentseverity <- total%>% select(UID,UKATEGORIE)%>% group_by(UKATEGORIE) %>% summarise(total.count=n()) %>% arrange(total.count)
by_Accidentseverity

#chart3
chartaccidentseverity<- ggplot(data=by_Accidentseverity, aes(x=UKATEGORIE, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Acccident_Severity")geom_text(aes(label=paste0(round(total.count, 2),"%")),position=position_dodge(width=0.9), vjust=-0.25)
chartaccidentseverity
```



Collision type 
```{r}
by_collisiontype <- total%>% select(UID,UART)%>% group_by(UART) %>% summarise(total.count=n()) %>% arrange(total.count)
by_collisiontype

#chart3
chartcollision<- ggplot(data=by_collisiontype, aes(x=UART, y=total.count, fill=UART))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Collision Type")+geom_text(aes(label=paste0(round(total.count, 2),"%")),position=position_dodge(width=0.9), vjust=-0.25)
chartcollision
```

accident type 
```{r}
by_accidenttype<- total%>% select(UID,UTYP1)%>% group_by(UTYP1) %>% summarise(total.count=n()) %>% arrange(total.count)
by_accidenttype

#chart3
chartaccidenttype<- ggplot(data=by_accidenttype, aes(x=UTYP1, y=total.count, fill=UTYP1))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Accident Type")
chartaccidenttype
```

 
recoding vaiables  

Road type  HIGHWAY 1- highway and 0 other roads 
```{r}
total <- total %>% 
  mutate(Highway= case_when(USTRKLASSE %in% 1:2 ~ 1,
                                      USTRKLASSE %in% 3:5 ~ 0))
```

```{r}
by_urbanhighway<- total%>% select(UID,UORTSLAGE,Highway)%>% group_by(UORTSLAGE,Highway) %>% summarise(total.count=n()) %>% arrange(total.count)
by_urbanhighway

#chart3
charturbanhighway<- ggplot(data=by_urbanhighway, aes(x=Highway, y=total.count, fill=UORTSLAGE))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Road type")+geom_text(aes(label=paste0(round(t_count, 2))), position = position_stack(vjust=0.9))+ scale_fill_discrete(name= "Area", labels=c("Urban", "Rural"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())+ scale_fill_manual(values =c("#00BFFF", "#1E90FF"))
charturbanhighway

```
  regroup in one variable callled AREA (1- highway, 3- rural, 2- urban)
  
```{r}
total <- total %>% 
  mutate(Area= case_when(Highway==1 & UORTSLAGE ==1 ~1,
                         Highway==1 & UORTSLAGE ==2 ~3,
                         Highway==0 & UORTSLAGE ==1 ~2,
                         Highway==0 & UORTSLAGE ==2 ~3
                         ))
```
```{r}
by_area<- total%>% select(UID,Area)%>% group_by(Area) %>% summarise(total.count=n()) %>% arrange(total.count)
by_area

#chart3
chartarea<- ggplot(data=by_area, aes(x=Area, y=total.count, fill=Area))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Area")
chartarea
```
  
  
area plus seve 

```{r}
areasev<- data %>% select(UID,UKATEGORIE, Area)%>% group_by(UKATEGORIE,Area) %>% summarise(t_count = n())%>% mutate(t_count = t_count/sum(t_count)*100)
areasev
chart_areasev<- ggplot(data=areasev, aes(x=UKATEGORIE, y=t_count, fill= Area))+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")), position = position_stack(vjust=0.9))+labs(y="Percentage", x= "Accident severity")+ scale_fill_discrete(name= "Area", labels=c("Higheway", "Urban", "Rural"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())#+scale_fill_brewer(palette = "Set2") #scale_fill_manual(values =c("#00BFFF", "#1E90FF", "#4169E1"))
chart_areasev
```


```{r}
total <- total %>% 
  mutate(Contributory_factor= case_when(BURSACHE1 %in% 1:11 ~1,
                         BURSACHE1 %in% 12:13 ~2,
                         BURSACHE1 %in% 14:15 ~3,
                         BURSACHE1 %in% 16:23 ~4,
                         BURSACHE1 %in% 24:33 ~1,
                         BURSACHE1 %in% 35:37 ~5,
                         BURSACHE1 %in% 38:42 ~1,
                         BURSACHE1 %in% 43:49 ~6,
                         BURSACHE1 %in% 50:55~7,
                         BURSACHE1 %in% 60:69 ~8
                         
                         
                         ))
```

```{r}
by_con_factors<- total%>% select(UID,Contributory_factor)%>% group_by(Contributory_factor) %>% summarise(total.count=n()) %>% arrange(total.count)
by_con_factors

#chart3
chartcon_factors<- ggplot(data=by_con_factors, aes(x=Contributory_factor, y=total.count, fill=Contributory_factor))+geom_bar(stat = "identity")+labs(y="Number of Accidents", x= "Contributory factors")
chartcon_factors
```

 
 
contibutorfactor  by sev 

```{r}
by_con_factorsEV<- total%>% select(UID,Contributory_factor, UKATAGEORIE )%>% group_by(Contributory_factor, UKATEGORIE) %>% summarise(total.count=n()) %>% arrange(total.count)
by_con_factorsEV

```

 
 speed limit
 
```{r}
#total<-total%>%
 # filter(UGESCHWBEGR %in% 005:130)


total <- total %>% 
  mutate(Speed_limit= case_when(UGESCHWBEGR %in% 005:030 ~"0-30",
                                UGESCHWBEGR %in% 040:060 ~"30-60",
                                UGESCHWBEGR %in% 070:100 ~"60-100",
                                UGESCHWBEGR %in% 110:130 ~"100-130",
                                UGESCHWBEGR %in% "Z07:z30" ~"a"
                         
                         
                         
                         ))
```
 
```{r}


by_spd<- total%>% select(UID,Speed_limit)%>% group_by(Speed_limit) %>% summarise(total.count=n()) %>% arrange(total.count)%>% drop_na(Speed_limit)
by_spd


```
 
```{r}
by_spd1<- total%>% select(UID,Speed_limit, UKATAGEORIE )%>% group_by(Speed_limit, UKATAGEORIE ) %>% summarise(total.count=n()) %>% arrange(total.count)%>% drop_na(Speed_limit)
by_spd1
```
 
 

Light conditions 8 1 - daylight, 0 - dark 

```{r}
total <- total %>%
  mutate(Light_cond=case_when(ULICHTVERH==0 ~ 1,
                              ULICHTVERH %in%1:2~0))
```


```{r}
bylight_cond<- total%>% select(UID,Light_cond)%>% group_by(Light_cond) %>% summarise(total.count=n()) %>% arrange(total.count)
bylight_cond

```

```{r}
bylight_cond1<- total%>% select(UID,Light_cond,UKATAGEORIE)%>% group_by(Light_cond,UKATAGEORIE) %>% summarise(total.count=n()) %>% arrange(total.count)
bylight_cond1
```

Road surafce condition 
```{r}
total <- total %>%
  mutate(Road_cond=case_when(USTRZUST1==0 ~ 1,
                              USTRZUST1 %in% 2:5~0))
```

number of particpants who were in an accident 
1 - dry 0- wet
```{r}
total <- total %>%
  mutate(N_of_participants=case_when(UANZBET==1~ 1,
                                     UANZBET==2~ 2,
                                     UANZBET==3~ 3,
                                     UANZBET %in% 4:30 ~ 4))
```


table road cond
```{r}
by_roadcond<- total%>% select(UID,Road_cond)%>% group_by(Road_cond) %>% summarise(total.count=n()) %>% arrange(total.count)
by_roadcond


```

road sev 
```{r}
by_roadcond1<- total%>% select(UID,Road_cond,UKATAGEORIE)%>% group_by(Road_cond, UKATAGEORIE) %>% summarise(total.count=n()) %>% arrange(total.count)
by_roadcond1

```



interaction terms where, who, when .... severity and the area 
```{r}
by_sevyear<- total%>% select(UID,UKATEGORIE, UJAHR)%>% group_by(UKATEGORIE, UJAHR) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevyear
chartsevyear<- ggplot(data = by_sevyear, aes(x=UJAHR, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Year")
chartsevyear
```


```{r}
by_sevarea<- total%>% select(UID,UKATEGORIE, Area)%>% group_by(UKATEGORIE, Area) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevarea
chartsevarea<- ggplot(data=by_sevarea, aes(x=Area, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Area")
chartsevarea
```
type of accidnet and collision 

```{r}
by_sevcol<- total%>% select(UID,UKATEGORIE, UART)%>% group_by(UKATEGORIE, UART) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevcol
chartsevcol<- ggplot(data=by_sevcol, aes(x=UART, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Collision Type")
chartsevcol
```


############Accidnet type analysis #############

accident type by severity 
```{r}
by_sevtype<- total%>% select(UID,UKATEGORIE, UTYP1)%>% group_by(UTYP1, UKATEGORIE) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevtype
chartsevtyp<- ggplot(data=by_sevtype, aes(x=UTYP1, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) + scale_fill_discrete(name= "Accident severity", labels=c("Slight","Serious", "Fatal"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartsevtyp

```

accident type by speed 
```{r}
by_typespeed<- total%>% select(UID,Speed_limit, UTYP1)%>% group_by(UTYP1,Speed_limit) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typespeed
chartspeedtype<- ggplot(data=by_typespeed, aes(x=UTYP1, y=total.count, fill=Speed_limit))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) +
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartspeedtype

```

accident type by area
```{r}
by_typearea<- total%>% select(UID,Area, UTYP1)%>% group_by(UTYP1,Area) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typearea
chartareatype<- ggplot(data=by_typearea, aes(x=UTYP1, y=total.count, fill=Area))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) + scale_fill_discrete(name="Area", labels=c("Highway","Urban", "Rural"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartareatype


```

pivot all
```{r}
by_typeall<- total%>% select(UID,Area,Speed_limit, UKATEGORIE, UTYP1)%>% group_by(UTYP1,Speed_limit, UKATEGORIE,Area) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typeall

```





subs set and filter 
```{r}
total1<- total
total1<- total1 %>%
  filter(BERSTZULKBA %in% 2010:2017 & UKATEGORIE%in% 1:2)


```

accident type by severity 
```{r}
by_sevtype1<- total%>% select(UID,UKATEGORIE, UTYP1)%>% group_by(UTYP1, UKATEGORIE) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevtype1
chartsevtyp1<- ggplot(data=by_sevtype1, aes(x=UTYP1, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) + scale_fill_discrete(name= "Accident severity", labels=c("Slight","Serious", "Fatal"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartsevtyp1

```

accident type by speed 
```{r}
by_typespeed1<- total1%>% select(UID,Speed_limit, UTYP1)%>% group_by(UTYP1,Speed_limit) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typespeed1
chartspeedtype1<- ggplot(data=by_typespeed1, aes(x=UTYP1, y=total.count, fill=Speed_limit))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) +
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartspeedtype1

```

accident type by area
```{r}
by_typearea1<- total1%>% select(UID,Area, UTYP1)%>% group_by(UTYP1,Area) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typearea1
chartareatype1<- ggplot(data=by_typearea1, aes(x=UTYP1, y=total.count, fill=Area))+geom_bar(stat = "identity")+labs(y="Percent %", x= "Accident Type")+geom_bar(stat = "identity")+geom_text(aes(label=paste0(round(t_count, 2), "%")),position=position_stack(vjust=0.5)) + scale_fill_discrete(name="Area", labels=c("Highway","Urban", "Rural"))+
  theme(legend.position="bottom", legend.direction="horizontal",
        legend.title = element_blank())
chartareatype1

```




pivot all
```{r}
by_typeall2<- total1%>% select(UID,Area,Speed_limit, UKATEGORIE, UTYP1)%>% group_by(UTYP1,Area,Speed_limit,UKATEGORIE) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_typeall2

```

















```{r}
cm<- cor(zip, method="pearson", use="everything")
corrplot(cm,tl.pos = "td", tl.cex = 0.5,type = "upper",diag = FALSE, order = "FPC")
```


```{r}
by_sevnum<- total%>% select(UID,N_of_participants, UKATEGORIE)%>% group_by(UKATEGORIE,N_of_participants ) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_sevnum
chartsevnum<- ggplot(data=by_sevnum, aes(x=N_of_participants, y=total.count, fill=UKATEGORIE))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Number of particpants ")
chartsevnum

```
```{r}
by_coltype<- total%>% select(UID,UART,UTYP1)%>% group_by(UART, UTYP1) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_coltype
chartcoltype<- ggplot(data=by_coltype, aes(x=UART, y=total.count, fill=UTYP1))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Collision vs accident type ")
chartcoltype
```

```{r}
by_colpart<- total%>% select(UID,UART,N_of_participants)%>% group_by(UART, N_of_participants) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_colpart
chartcolpart<- ggplot(data=by_colpart, aes(x=UART, y=total.count, fill=N_of_participants))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Collision vs Number of Accidnets ")
chartcolpart
```

```{r}
by_colarea<- total%>% select(UID,UART,Area)%>% group_by(UART, Area) %>% summarise(total.count=n()) %>% arrange(total.count)%>% mutate(total.count = total.count/sum(total.count)*100)
by_colarea
chartcolarea<- ggplot(data=by_colarea, aes(x=UART, y=total.count, fill=Area))+geom_bar(stat = "identity")+labs(y="Percentage of accidents", x= "Collision vs Area ")
chartcolarea
```
Write the file to reuse later 

```{r}
write.csv(total, "C:/Forscher/3891-2019/Arbeitsverzeichnis/descr_final.csv", sep = "," )
```


###########---------------------Next Filter 

Rename the manufacture date of vehicle 

```{r}

```

filter vehicles made from 2010 onwards 
